Review Paper Draft

### 2020-06-25 main.tex
Edits to Section 4 and 5

%------------------------------%
\subsubsection{LMM-Lasso variants}

@@

\pb{should this paragraph move to next section?}
\anna{yes} Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. \pb{So, for example, $\eta=0.5$ and $xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).

@@

%------------------------------%
\subsection{PC-Lasso and LMM-Lasso outperform Lasso}
%------------------------------%
Firstly, these penalized regression methods were compared in the presence of varying levels of environmental confounding. Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.  Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of figure \ref{fig:mse} shows that when $\xi = 0$, and no environmental confounding is present, lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that the top 10 principal components do not explain any variability in the outcome, yet are included in the final model, since they are unpenalized. The bottom row of figure \ref{fig:mse} illustrates the resulting MSE when $\xi = 0.8$ and thus 40\% of the variability of the phenotype is attributable to environmental confounding. In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso. This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.

@@

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when simulated subpopulations are small}
%------------------------------%
Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms. Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data. We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.

@@

@@

Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. \cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.

@@

@@

In section 3 briefly mentioned the existence of to LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. The LMM-Lasso results shown so far are those of LMM-Lasso-Rakitsch, though they are representative of LMM-Lasso-ggmix. The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods which aim to correct for population structure and environmental heterogeneity. However, we also wish to briefly discuss the key differences between methods, and provide guidance for potential users of either method.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion} \label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this work, we have reviewed the concept of population stratification and defined explicitly the non-genetic, environmental mechanism by which it may contribute to confounding the genotype-phenotype relationship. We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences. In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.

@@

>>>>

%------------------------------%
\subsubsection{LMM-Lasso variants}
\label{sec:rak-ggmix}

@@

\pb{should this paragraph move to next section?}
\anna{yes} Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. \pb{So, for example, $\eta=0.5$ and $\xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).

@@

%------------------------------%
\subsection{PC-Lasso and LMM-Lasso outperform Lasso in the presence of environmental confounding}
%------------------------------%

We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied. Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.  Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome. The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$ (i.e., 40\% of the variability in the phenotype is attributable to environmental confounding). In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso. This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.

\pb{Need to describe how we select $\lambda$}

@@

\pb{Just FYI, it's hard to tell bias and var apart in black and white}

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when simulated subpopulations are small}
%------------------------------%

\pb{I find ``small'' subpopulations unclear; how do you feel about ``fine'' and ``coarse'' structure?}

Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms. Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data. We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.

\pb{should describe our empirical structure here -- is it fine or coarse?}

@@

\pb{these are nice figures, but to me, better for supplement}

@@

Phenotypes for these simulations were again simulated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.

@@

\pb{We should discuss this section; to me, this section comes across as overly complicated and the point is not really coming across}

@@

Section~\ref{sec:rak-ggmix} briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. As we will see, the two generally perform quite similarly. Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of either LMM-Lasso method.

The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding. However, given the generally strong performance of LMM-Lasso, we also wish to briefly discuss the key differences between these two ways of implementing it, and provide guidance for potential users.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion} \label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this work, we have reviewed the concept of population stratification and defined explicitly the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship. We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences. In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.

@@

\pb{Introduce penalizedLMM here, or in next paper?}
